Nonparametric selective bootstrap

### 2015-05-22 notes/nonparam.tex
a little more cleanup on nonparametric notes

We will apply $\hat{M}$ on data of the form
$$
\hat{Z}_n = \frac{1}{n} \sum_{i=1}^n Z(t_i) 
$$
where $t_i \overset{\text{IID}}{\sim} F$ where $F$
is some distribution on some sample space ${\cal X}$. In the pairs bootstrap for 
regression context, we should think of $t_i=(y_i, X_i)$ and $F$ 
is a distribution on $\real \times \real^p$ and, if we use
forward stepwise our
selection is a function of $\frac{1}{n}(X^Ty, X^TX)$
and our $Z$ should be taken to be
$$
Z(t_i) = (y_iX_i, X_iX_i^T).
$$

Note that $Z$ is a statistical functional that we might write as
$$
Z(F) = E_{F}(Z(\tilde{y}, \tilde{X}))
$$
and what we really observe $\hat{Z}_n = Z(\hat{F}_n)$.

@@

{\bf We assume that $Z$ is measurable with respect to $\gamma$, i.e.
$\gamma$ includes $Z$ as part of its coordinates or something. This means
that it makes sense to talk about $\hat{M}$ evaluated
on a sample of $\gamma$'s.}

@@

Suppose we have a one-sample problem and if the empirical skewness
is larger than some threshold, we test whether the median of the 
distribution, otherwise we report the mean. In one case,
$\gamma_M$ would compute sample skewness and sample median, in the other
it would compute the sample skewness and sample mean. When choosing to test the sample mean, I suppose the model is all distributions on $\real$ for
which the sample mean and sample skewness and sample mean 
are asymptotically Gaussian. When
reporting the sample median, the model is all distributions on $\real$
for which the sample skewness and sample median are asymptotically jointly Gaussian.

@@

We have chosen $\eta_M$ based on the outcome $\hat{M}(\hat{Z}_n)$. Following the general setup of the
selected model framework, we just have to design
a test for a fixed $M$.

For large enough $n$, under our model $M$ suppose that
$$
\gamma(\hat{F}_n) \approx N(\gamma(F), n^{-1} \Sigma(F))
$$
where, for some distribution $G$ on $\real \times \real^p$
$$
\Sigma(F) = E_{F}\left((\gamma(\tilde{t}) - \gamma(F))(\gamma(\tilde{t}) - \gamma(F))^T \right)
$$
where $\tilde{t}$ is a sample of size 1 from $F$.

@@

If this is our reference distribution, we should base our distribution on the 
conditional law 
$$\Sigma(F)^{-1}\gamma(\hat{F}_n) \big \vert P_{\eta}^{\perp}\Sigma(F)^{-1}\gamma(\hat{F}_n).$$ 
where $P_{\eta}=\eta\eta^{\dagger}$ and $P_{\eta}^{\perp} = I - P_{\eta}$.
This is because
we know that, under $H_0:\eta^T\gamma(F)=0$
this distribution does not depend on $P_{\eta}^{\perp}\gamma(F)$ which
is everything except what we wanted to test.

This 
distribution has covariance $n^{-1} \Theta(\eta, \Sigma(F))$ where
$$
\Theta(\eta, \Sigma(F)) = \left(\Sigma(F)^{-1} - \Sigma(F)^{-1} P_{\eta}^{\perp} \left(P_{\eta}^{\perp} \Sigma(F)^{-1} P_{\eta}^{\perp} \right)^{\dagger}  P_{\eta}^{\perp} \Sigma(F)^{-1}  \right).
$$
Under $H_0$, this distribution has mean 0.

If $W \sim N(\gamma(F), n^{-1}\Sigma(F))$, then, under $H_0$
this distribution can be realized by 
$$
L(\eta, \Sigma(F), W) = \Sigma(F)^{-1} W - \Sigma(F)^{-1} P_{\eta}^{\perp} \left(P_{\eta}^{\perp} 
\Sigma(F)^{-1} P_{\eta}^{\perp} \right)^{\dagger}  P_{\eta}^{\perp} W.
$$

In summary, if $W \sim N(\gamma(F), n^{-1}\Sigma(F))$ 
then under $H_0: \eta^T\gamma(F)=0$ we have
$$
L(\eta, \Sigma(F), W) \sim N(0, n^{-1} \Theta(\eta, \Sigma(F))).
$$

Therefore, if we can
sample from this distribution, we know how to
do a parametric unselective test of $H_0:\eta_M^T\gamma(F)=0$.

@@

@@

If $W \sim N(\mu, \Sigma_M(F))$ is any normal distribution
such that $\eta^T\mu=0$ then by familiar exponential family principles, the above law
is equal to the law of
$$
L_M(\eta_M, \Sigma_M(F), W) \big \vert \hat{M}(W) = M.
$$

{\bf This means it is enough to find any normal distribution
$W \sim N(\mu, \Sigma_M(F))$ such that $\eta^T\mu=0$ to carry out the
selective test.
}

@@

Therefore, our unselective bootstrap should be giving samples from
roughly $ N(\gamma(\hat{F}_n), n^{-1} \Sigma(\hat{F}_n))$. That is,
taking $B$ very large:
$$
\frac{1}{B}\sum_{b=1}^B \eta^T \gamma(\hat{F}^{*,b}_n) \approx \eta^T\gamma(\hat{F}_n).
$$
This suggests we correct the unselective bootstrap by drawing
$$
\gamma(\hat{F}^{*,b}_n) - P_{\eta}\gamma(\hat{F}_n) | \hat{F}_n \approx N(P_{\eta}^{\perp}\gamma(\hat{F}_n) ,n^{-1} \Sigma(\hat{F}_n))
$$
and we see that
$$
L(\eta, \Sigma(\hat{F}_n), \gamma(\hat{F}^{*,b}_n) - P_{\eta} \gamma(\hat{F}_n)) \big \vert \hat{F}_n
 \approx N(0 ,n^{-1} \Theta(\eta, \Sigma(\hat{F}_n))).
$$
As $n$ gets larger, $\Sigma(\hat{F}_n)$ will approach $\Sigma(F)$. 
If the RHS above were actually $$N(0 ,n^{-1} \Theta(\eta, \Sigma(\hat{F}_n))),$$ then
by our earlier observation at the end of Section \ref{sec:selective} this is sufficient to carry out the selective test.
However, the variance is not exactly correct. In general,
plugging in consistent estimates of variance is OK in selective inference (not sure
if Xiaoying and I put it in the first paper on asymptotics, but definitely in the
randomized setting we will have a concrete statement about this). In Rob's example of $n=30$, the sample size might be too small still, but asymptotically
I think it will be fine. Maybe a degrees of freedom correction will help.

So,
we draw nonparametric bootstrap samples
$$
\gamma(\hat{F}^{*,b}_n) - P_{\eta}\gamma(\hat{F}_n)
$$
keeping only those for which $\hat{M}(\gamma(\hat{F}^{*,b}_n) - P_{\eta}\gamma(\hat{F}_n)) = \hat{M}(\gamma(\hat{F}_n))$.

For those that survive, we compute
$$
L(\eta, \Sigma(\hat{F}_n), \gamma(\hat{F}^{*,b}_n) - P_{\eta}\gamma(\hat{F}_n))
$$
and take a test statistic of our choosing.

>>>>

We will apply $\hat{M}$ on data of the form
$$
\hat{Z}_n = \frac{1}{n} \sum_{i=1}^n Z(t_i) 
$$
where $t_i \overset{\text{IID}}{\sim} F$ where $F$
is some distribution on some sample space ${\cal X}$. In the pairs bootstrap for 
regression context, we should think of $t_i=(y_i, X_i)$ and $F$ 
is a distribution on $\real \times \real^p$ and, if we use
forward stepwise our
selection is a function of 
$$\frac{1}{n}(X^Ty, X^TX)$$
and our $Z$ should be taken to be
$$
Z(t_i) = (y_iX_i, X_iX_i^T).
$$

Note that $Z$ is a statistical functional that we might write as
$$
Z(F) = E_{F}(Z(t))
$$
and what we really observe $\hat{Z}_n = Z(\hat{F}_n)$.

@@

We assume that $Z$ is measurable with respect to $\gamma$, i.e.
$\gamma$ includes $Z$ as part of its coordinates or something. This means
that it makes sense to talk about $\hat{M}$ evaluated
on a sample of $\gamma$'s.

@@

Here is an ``example.''
Suppose we have a one-sample problem and if the empirical skewness
is larger than some threshold, we test whether the median of the 
distribution, otherwise we report the mean. In one case,
$\gamma_M$ would compute sample skewness and sample median, in the other
it would compute the sample skewness and sample mean. When choosing to test the sample mean, I suppose the model is all distributions on $\real$ for
which the sample mean and sample skewness and sample mean 
are asymptotically Gaussian. When
reporting the sample median, the model is all distributions on $\real$
for which the sample skewness and sample median are asymptotically jointly Gaussian.

@@

For large enough $n$, under our model $M$ suppose that
$$
\gamma(\hat{F}_n) \approx N(\gamma(F), n^{-1} \Sigma(F))
$$
where
$$
\Sigma(F) = E_{F}\left((\gamma(t) - \gamma(F))(\gamma(t) - \gamma(F))^T \right).
$$

@@

If this is our reference distribution, we should base our test on the 
conditional law 
$$n^{1/2}\Sigma(F)^{-1}\gamma(\hat{F}_n) \big \vert n^{1/2} P_{\eta}^{\perp}\Sigma(F)^{-1}\gamma(\hat{F}_n).$$ 
where $P_{\eta}=\eta\eta^{\dagger}$ and $P_{\eta}^{\perp} = I - P_{\eta}$.
This is because
we know that, under $H_0:\eta^T\gamma(F)=0$
this distribution does not depend on $P_{\eta}^{\perp}\gamma(F)$ which
is everything except what we wanted to test.

This 
law has covariance $n^{-1} \Theta(\eta, \Sigma(F))$ where
$$
\Theta(\eta, \Sigma(F)) = \left(\Sigma(F)^{-1} - \Sigma(F)^{-1} P_{\eta}^{\perp} \left(P_{\eta}^{\perp} \Sigma(F)^{-1} P_{\eta}^{\perp} \right)^{\dagger}  P_{\eta}^{\perp} \Sigma(F)^{-1}  \right)
$$
which is an $O(1)$ covariance matrix.
Under $H_0$, this law has mean 0. Under any alternative, its mean grows like $n^{1/2}$.

Under $H_0$
this distribution can be realized by 
$$
L(\eta, \Sigma(F), W) = n^{1/2} \left(\Sigma(F)^{-1} W - \Sigma(F)^{-1} P_{\eta}^{\perp} \left(P_{\eta}^{\perp} 
\Sigma(F)^{-1} P_{\eta}^{\perp} \right)^{\dagger}  P_{\eta}^{\perp} W \right)
$$
where $W \sim N(\mu, n^{-1}\Sigma(F))$, is any normal distribution with
$\eta^T\mu=0$.

In other words, if $W \sim N(\mu, n^{-1}\Sigma(F))$ 
then under $H_0: \eta^T\gamma(F)=0$ we have, asymptotically
$$
 L(\eta, \Sigma(F), \gamma(\hat{F}_n))  \overset{n \rightarrow \infty}{\Rightarrow} L(\eta, \Sigma(F), W) = N(0,  \Theta(\eta, \Sigma(F))).
$$

Therefore, if we can
sample from this distribution, we know how to
do an asymptotic unselective test of $H_0:\eta_M^T\gamma(F)=0$ using the 
normal approximation.

@@

Recall that we have chosen $\eta_M$ based on the outcome $\hat{M}(\hat{Z}_n)$. Following the general setup of the
selected model framework, we just have to design
a test for a fixed $M$.

@@

Now, if $W \sim N(\mu, n^{-1}\Sigma(F))$ is any normal distribution
such that $\eta^T\mu=0$ then by familiar exponential family principles, the above law
is equal to the law of
$$
L_M(\eta_M, n^{-1}\Sigma(F), W) \big \vert \hat{M}(W) = M.
$$

{\bf This means it is enough to find any normal distribution
$W \sim N(\mu, n^{-1}\Sigma(F))$ such that $\eta^T\mu=0$ to carry out the
selective test.
}

@@

Therefore, our unselective bootstrap should be giving samples from
roughly $ N(\gamma(\hat{F}_n), n^{-1} \Sigma(\hat{F}_n))$. That is,
taking $B$ very large:
$$
\frac{1}{B}\sum_{b=1}^B \eta^T \gamma(\hat{F}^{*,b}_n) \approx \eta^T\gamma(\hat{F}_n).
$$
This suggests we correct the unselective bootstrap by drawing
$$
\gamma(\hat{F}^{*,b}_n) - P_{\eta}\gamma(\hat{F}_n) | \hat{F}_n \approx N(P_{\eta}^{\perp}\gamma(\hat{F}_n) ,n^{-1} \Sigma(\hat{F}_n)).
$$
With this correction, we see that
$$
L(\eta, \Sigma(\hat{F}_n), \gamma(\hat{F}^{*,b}_n) - P_{\eta} \gamma(\hat{F}_n)) \big \vert \hat{F}_n
 \approx N(0 ,n^{-1} \Theta(\eta, \Sigma(\hat{F}_n))).
$$
As $n$ gets larger, $\Sigma(\hat{F}_n)$ will approach $\Sigma(F)$. 
If the RHS above were actually $$N(0 ,n^{-1} \Theta(\eta, \Sigma(F))),$$ then
by our earlier observation at the end of Section \ref{sec:selective} this is sufficient to carry out the selective test exactly.
However, the variance is not exactly correct. In general,
plugging in consistent estimates of variance is OK in selective inference (not sure
if Xiaoying and I put it in the first paper on asymptotics, but she has given me a proof and definitely in the
randomized paper we will have a concrete statement about this). In Rob's example of $n=30$, the sample size might be too small still, but asymptotically
I think it will be fine. Maybe a degrees of freedom correction will help.

The final algorithm is:
we draw nonparametric bootstrap samples
$$
\gamma(\hat{F}^{*,b}_n) - P_{\eta}\gamma(\hat{F}_n)
$$
keeping only those for which $\hat{M}(\gamma(\hat{F}^{*,b}_n) - P_{\eta}\gamma(\hat{F}_n)) = \hat{M}(\gamma(\hat{F}_n))$.

For those that survive, we compute
$$
L(\eta, \Sigma(\hat{F}_n), \gamma(\hat{F}^{*,b}_n) - P_{\eta}\gamma(\hat{F}_n))
$$
and take any test statistic of our choosing.
